Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-09-25 article.tex
Section 4 updates

\paragraph{Case 6: Vibration of a 3-D Cantilevered Beam}
This 3-D, dynamic, finite strain case geometry consists of a $2 \times 0.2 \times 0.2$ m cuboid column and was proposed in its initial form by \citet{Tukovic2007}.
%(Figure \ref{fig:dynamic_cantilever})
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%\begin{eqnarray}
%	|\bb{T}| = \frac{\mathcal{F} E I}{A L^2} \text{ Pa}
%\end{eqnarray}
A sudden, constant traction $\bb{T} = \left(0.1, 0.1, 0 \right)$ MPa is applied to the upper surface.
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%In the current study, $\mathcal{F} = 1$.
A neo-Hookean hyperelastic material is assumed with $E = 15.293$ MPa, $\nu = 0.3$ and density $\rho = 1000$ kg m$^{-3}$, while the area $A = (0.2)(0.2) = 0.04$ m$^2$ and the second moment of area $I = \frac{1}{7\,500}$ m$^4$. %0.0001333
The geometric and material parameters were chosen for the first natural frequency to be 1 Hz in the small strain limit.
The magnitude of the applied traction was chosen to ensure significant geometric nonlinearity in the case with the upper surface of the beam expected to drop below the plane of the lower surface.
Four succesively refined hexahedral meshes are generated using the OpenFOAM \texttt{blockMesh} utility, with cell counts of 270, $2\,160$, $17\,280$ and $138\,240$.
%A fixed time step size of 0.001617274174 s is chosen, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
%A fixed time step size of 1 ms is chosen.
The time step size is 1 ms, and the total period is 10 s.
% corresponding to 10 expected oscillation periods.
%, corresponding to a Courant number of six on the mesh with $2\,160$ cells.

The deformed configuration of the beam at six time steps is shown in Figure \ref{fig:dynamic_cantilever}(a) for the mesh with $2\,160$ cells, where the upper surface is seen to drop below the horizontal from approximately t = 0.2 s to 0.32 s.
The corresponding displacement magnitude of the centre of the upper surface of the beam vs time is shown in Figure \ref{fig:dynamic_cantilever}(b).
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.5\textwidth]{figures/cantilever} 
	\subfigure[Displacement magnitude for t = $\left\{0, \, 0.1, \, 0.2\right\}$ s (translucent) and t = $\left\{0.3, \, 0.4, \, 0.5 \right\}$ s (opaque) for the mesh with $138\,240$ cells]
	{
	   \includegraphics[width=0.4\textwidth]{figures/cantilever_deformed}
   	}
	\subfigure[Displacement magnitude of the centre of the upper surface of the beam vs time]
	{
	   \includegraphics[width=0.5\textwidth]{figures/cantilever_disp} 
   	   %\includegraphics[height=0.45\textwidth]{figures/cantilever_sigmaEq}  
   	}
   \caption{Deflection of a 3-D Cantilevered Beam}
   \label{fig:dynamic_cantilever}
\end{figure}
\hl{It would be good to add a reference results, e.g. from FE}

@@

\paragraph{Case 8: Elastic plate behind a rigid cylinder}
\citet{Turek2006} proposed this well-known fluid-solid interaction benchmark case as an extension of the classic flow around a cylinder in a channel problem \cite{Ferziger2002}.
The \texttt{FSI3} variant of the case examined here (Figure \ref{fig:hronTurek-geometry} consists of a horizontal channel (0.41 m in height, 2.5 m in length) with a rigid cylinder of radius 0.05 m, where the cylinder centre is 0.2 m from the bottom and inlet (left) boundaries.
A parabolic velocity is prescribed at the inlet velocity with a mean value of 2 m/s.
A St.\ Venant-Kirchhoff hyperelastic plate ($E = 5.6$ MPa, $\nu = 0.4$) of 0.35 m in length and 0.02 m in height is attached to the right-hand side of the rigid cylinder.
The fluid model is assumed to be isothermal, incompressible, and laminar and adopts the segregated PIMPLE solution algorithm.
The fluid's kinematic viscosity is 0.001 m$^2$/s and density is 1000 kg/m$^3$, while the solid's density is $10\,000$ kg/m$^3$.
The interface- quasi-Newton with inverse Jacobian from a least-squares model coupling approach \cite{Degroote2009} is employed for the fluid-solid interaction coupling.
The fluid-solid interface residual is reduced by four orders of magnitude within each time step.
Further details of the fluid-solid interaction procedure are found in \citet{Tukovic2018a}.
\begin{figure}[htbp]
   \centering
%	\subfigure[Geometry (taken from \citet{Tukovic2018a})]
%	{
	   \includegraphics[width=0.7\textwidth]{figures/hronTurek-geometry} 
%   	}
   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
   \label{fig:hronTurek-geometry}
\end{figure}

Three successively refined quadrilateral meshes are employed:
The fluid region meshes have $1\,252$, $5\,008$, and $20\,032$ cells, while the solid region meshes have 156,  624 and $2\,496$ cells.
The total simulation time is \hl{6 or 20} s, and the time-step size is 0.5 ms. 
\begin{figure}[htbp]
   \centering
	\subfigure[Mesh containing $5\,336$ cells in the fluid region and 630 cells in the solid region, where a closeup of the plate is shown in the right]
	{
   		\includegraphics[width=0.7\textwidth]{figures/hronTurek-mesh}
	}
	\subfigure[Close-up of the mesh around the plate]
	{
   		\includegraphics[width=0.4\textwidth]{figures/hronTurek-mesh-zoom}  
   	}
   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
   \label{fig:hronTurek-mesh}
\end{figure}

Figure \ref{fig:hronTurek-results} shows the velocity field in the fluid region and displacement magnitude in the solid region at $t = 4.42$ s, corresponding to a peak in the vertical displacement of the plate free-end oscillation.
The fluid is seen to accelerate as it passes the cylinder, with regular vortices being thrown off the plate, causing it to oscillate.
\begin{figure}[htbp]
   \centering
	   \includegraphics[width=0.8\textwidth]{figures/hronTurek-results} 
%	\subfigure[Velocity field in the fluid region and displacement magnitude in the solid region at $t = 4.42$ s]
%	{
%	   \includegraphics[width=0.8\textwidth]{figures/hronTurek-results} 
%   	}
%	\subfigure[Vertical displacement of the end of the plate vs time]
%	{
%   		\includegraphics[width=0.3\textwidth]{figures/placeholder}  
%   	}
   \caption{Elastic plate behind a rigid cylinder case results for the mesh with $5\,336$ cells in the fluid region and 630 in the solid region}
   \label{fig:hronTurek-results}
\end{figure}
The predicted mean, amplitude and frequency of the vertical and horizontal displacement of the end of the plate are compared with the results from \citet{Turek2006} in Table \ref{tab:hronTurekDisp}.
As directed in \citet{Turek2006}, the maximum and minimum values for the last period are used to calculate the mean and amplitude:
\begin{eqnarray}
	\text{mean} = \frac{1}{2}(\text{max} + \text{min}) \\
	\text{amplitude} = \frac{1}{2}(\text{max} - \text{min})
\end{eqnarray}
while the frequency is calculated as the inverse of the period.
\begin{table}[htb]
	\centering
		\begin{tabular}{lll}
			\hline
			Mesh (number of fluid and solid cells) & $u_x$ (in mm) & $u_y$ (in mm) \\
			\hline
			1 ($1\,252$ + 156)  & $-0.1 \pm 0.01 \,[ 18.5 ]$ & $1.60 \pm 0.20 \,[ 5.5 ]$ \\
			2 ($5\,008$ + 624)  & $-2.13 \pm 2.07 \,[ 11 ]$ & $2.00 \pm 28.84 \,[ 5.5 ]$ \\
			3 ($20\,032$ + $2\,496$)  & $-2.62 \pm 2.46 \,[11]$ & $1.80 \pm 32.01 \,[ 5.5 ]$ \\
			\hline
			\citet{Turek2006} & $-2.69 \pm 2.53\,[10.9]$ & $1.48 \pm 34.38\,[5.3]$ \\
			\hline
		\end{tabular}
	\caption{Average predicted displacements (mean $\pm$ amplitude $[$frequency$]$) of the free end of the plate. \hl{This data is taken as the average from 4-6 seconds: I am now running the case up to 20 seconds as per the benchmark}}
	\label{tab:hronTurekDisp}
\end{table}

>>>>

\paragraph{Case 6: Vibration of a 3-D Cantilevered Beam}
This 3-D, dynamic, finite strain case geometry consists of a $2 \times 0.2 \times 0.2$ m cuboid column and was proposed in its initial form by \citet{Tukovic2007}.
%(Figure \ref{fig:dynamic_cantilever})
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%\begin{eqnarray}
%	|\bb{T}| = \frac{\mathcal{F} E I}{A L^2} \text{ Pa}
%\end{eqnarray}
A sudden, constant traction $\bb{T} = \left(0.1, 0.1, 0 \right)$ MPa is applied to the upper surface.
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%In the current study, $\mathcal{F} = 1$.
A neo-Hookean hyperelastic material is assumed with $E = 15.293$ MPa, $\nu = 0.3$ and density $\rho = 1000$ kg m$^{-3}$, while the area $A = 0.04$ m$^2$ and the second moment of area $I = \frac{1}{7\,500}$ m$^4$. %0.0001333
The geometric and material parameters were chosen for the first natural frequency to be 1 Hz in the small strain limit.
The magnitude of the applied traction was chosen to ensure significant geometric nonlinearity with the upper surface of the beam expected to drop below the plane of the lower surface.
Four succesively refined hexahedral meshes are generated using the OpenFOAM \texttt{blockMesh} utility, with cell counts of 270, $2\,160$, $17\,280$ and $138\,240$.
%A fixed time step size of 0.001617274174 s is chosen, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
%A fixed time step size of 1 ms is chosen.
The time step size is 1 ms, and the total period is 10 s.
% corresponding to 10 expected oscillation periods.
%, corresponding to a Courant number of six on the mesh with $2\,160$ cells.

The deformed configuration of the beam at six time steps is shown in Figure \ref{fig:dynamic_cantilever}(a) for the mesh with $2\,160$ cells, where the upper surface is seen to drop below the horizontal from approximately t = 0.2 s to 0.32 s.
The corresponding displacement magnitude of the centre of the upper surface of the beam vs time is shown in Figure \ref{fig:dynamic_cantilever}(b).
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.5\textwidth]{figures/cantilever} 
	\subfigure[Displacement magnitude for t = $\left\{0, \, 0.1, \, 0.2\right\}$ s (translucent) and t = $\left\{0.3, \, 0.4, \, 0.5 \right\}$ s (opaque) for the mesh with $138\,240$ cells]
	{
	   \includegraphics[width=0.4\textwidth]{figures/cantilever_deformed}
   	} \quad
	\subfigure[Displacement magnitude of the centre of the upper surface of the beam vs time]
	{
	   \includegraphics[width=0.5\textwidth]{figures/cantilever_disp} 
   	   %\includegraphics[height=0.45\textwidth]{figures/cantilever_sigmaEq}  
   	}
   \caption{Deflection of a 3-D Cantilevered Beam}
   \label{fig:dynamic_cantilever}
\end{figure}
\hl{It would be good to add a reference results, e.g. from FE}

@@

\paragraph{Case 8: Elastic plate behind a rigid cylinder}
\citet{Turek2006} proposed this well-known fluid-solid interaction benchmark case as an extension of the classic flow around a cylinder in a channel problem \cite{Ferziger2002}.
The \texttt{FSI3} variant of the case examined here (Figure \ref{fig:hronTurek-mesh}) consists of a horizontal channel (0.41 m in height, 2.5 m in length) with a rigid cylinder of radius 0.05 m, where the cylinder centre is 0.2 m from the bottom and inlet (left) boundaries.
A parabolic velocity is prescribed at the inlet velocity with a mean value of 2 m/s.
A St.\ Venant-Kirchhoff hyperelastic plate ($E = 5.6$ MPa, $\nu = 0.4$) of 0.35 m in length and 0.02 m in height is attached to the right-hand side of the rigid cylinder.
The fluid model is assumed to be isothermal, incompressible, and laminar and adopts the segregated PIMPLE solution algorithm.
The fluid's kinematic viscosity is 0.001 m$^2$/s and density is 1000 kg/m$^3$, while the solid's density is $10\,000$ kg/m$^3$.
The interface- quasi-Newton with inverse Jacobian from a least-squares model coupling approach \cite{Degroote2009} is employed for the fluid-solid interaction coupling.
The fluid-solid interface residual is reduced by four orders of magnitude within each time step.
Further details of the fluid-solid interaction procedure are found in \citet{Tukovic2018a}.
%\begin{figure}[htbp]
%   \centering
%%	\subfigure[Geometry (taken from \citet{Tukovic2018a})]
%%	{
%	   \includegraphics[width=0.7\textwidth]{figures/hronTurek-geometry} 
%%   	}
%   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
%   \label{fig:hronTurek-geometry}
%\end{figure}
Three successively refined quadrilateral meshes are employed:
The fluid region meshes have $1\,252$, $5\,008$, and $20\,032$ cells, while the solid region meshes have 156,  624 and $2\,496$ cells.
The total simulation time is 20 s, and the time-step size is 0.5 ms. 
\begin{figure}[htbp]
   \centering
	\subfigure[Mesh containing $5\,336$ cells in the fluid region and 630 cells in the solid region, where a closeup of the plate is shown in the right]
	{
   		\includegraphics[width=0.7\textwidth]{figures/hronTurek-mesh}
	}
	\subfigure[Close-up of the mesh around the plate]
	{
   		\includegraphics[width=0.4\textwidth]{figures/hronTurek-mesh-zoom}  
   	}
   \caption{Elastic plate behind a rigid cylinder case geometry and mesh}
   \label{fig:hronTurek-mesh}
\end{figure}

Figure \ref{fig:hronTurek-results} shows the velocity field in the fluid region and displacement magnitude in the solid region at $t = 4.42$ s, corresponding to a peak in the vertical displacement of the plate free-end oscillation.
The fluid is seen to accelerate as it passes the cylinder, with regular vortices being thrown off the plate, causing it to oscillate.
\begin{figure}[htbp]
   \centering
	   \includegraphics[width=0.8\textwidth]{figures/hronTurek-results} 
%	\subfigure[Velocity field in the fluid region and displacement magnitude in the solid region at $t = 4.42$ s]
%	{
%	   \includegraphics[width=0.8\textwidth]{figures/hronTurek-results} 
%   	}
%	\subfigure[Vertical displacement of the end of the plate vs time]
%	{
%   		\includegraphics[width=0.3\textwidth]{figures/placeholder}  
%   	}
   \caption{Elastic plate behind a rigid cylinder case results for the mesh with $5\,336$ cells in the fluid region and 630 in the solid region. The velocity magnitude is shown in the fluid region while the displacement magnitude is shown in the solid region.}
   \label{fig:hronTurek-results}
\end{figure}
The predicted mean, amplitude and frequency of the vertical and horizontal displacement of the end of the plate are compared with the results from \citet{Turek2006} in Table \ref{tab:hronTurekDisp}.
As directed in \citet{Turek2006}, the maximum and minimum values for the last period are used to calculate the mean and amplitude:
\begin{eqnarray}
	\text{mean} = \frac{1}{2}(\text{max} + \text{min}), \quad\quad
	\text{amplitude} = \frac{1}{2}(\text{max} - \text{min})
\end{eqnarray}
while the frequency is calculated as the inverse of the period.
\begin{table}[htb]
	\centering
		\begin{tabular}{lll}
			\hline
			Mesh (number of fluid and solid cells) & $u_x$ (in mm) & $u_y$ (in mm) \\
			\hline
			1 ($1\,252$ + 156)  & $-0.1 \pm 0.01 \,[ 18.5 ]$ & $1.60 \pm 0.20 \,[ 5.5 ]$ \\
			2 ($5\,008$ + 624)  & $-2.13 \pm 2.07 \,[ 11 ]$ & $2.00 \pm 28.84 \,[ 5.5 ]$ \\
			3 ($20\,032$ + $2\,496$)  & $-2.62 \pm 2.46 \,[11]$ & $1.80 \pm 32.01 \,[ 5.5 ]$ \\
			\hline
			\citet{Turek2006} & $-2.69 \pm 2.53\,[10.9]$ & $1.48 \pm 34.38\,[5.3]$ \\
			\hline
		\end{tabular}
	\caption{Average predicted displacements (mean $\pm$ amplitude $[$frequency$]$) of the free end of the plate. \hl{This data is taken as the average from 4-6 seconds: I am now running the case up to 20 seconds as per the benchmark}}
	\label{tab:hronTurekDisp}
\end{table}
